---
title: "A3: Incarceration"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(scales)
library(maps)

data_prison <- read.csv("https://github.com/melaniewalsh/Neat-Datasets/blob/main/us-prison-pop.csv?raw=true")

source('chart1.R')
source('chart2.R')
source('map.R')

```

### Introduction

I am choosing to evaluate the difference between the black female and white female prison population. Because America is the country with the greatest number of people incarcerated in the world and Black people are incarcerated at disproportionately high rates, I wanted to analyze this disproportion within the female population in prison. 

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

These will likely be calculated using your DPLYR skills, answering questions such as:

-   What is the average value of my variable across all the counties (in the current year)?
-   Where is my variable the highest / lowest?
-   How much has my variable change over the last N years?

Feel free to calculate and report values that you find relevant. Again, remember that the purpose is to think about how these measure of incarceration vary by race.

### The Dataset

The data was released by the Vera Institute of Justice  They employed two Bureau of Justice Statistics (BJS) data collections: the Census of Jails (COJ) and the Annual Survey of Jails (ASJ). Furthermore, this project was funded by the Robert W. Wilson Charitable Trust. The data was collected to fill the gap of data needed to analyze the trends of causes, consequences, and inequalities. Most of my analysis used the data from "Dataset about prison population per county/state." There is 153811 observations (rows) and 36 features (columns). A possible limitation of the data set is that there are a lot of NA. There are also large gaps between the time taken between the data. A question of power that is worth considering when working with this data set is whether some of the NA in the data set is the result of hidding the statistics in order to conceal inequalities. 

# Who collected the data?

# How was the data collected or generated?

# Why was the data collected?

# How many observations (rows) are in your data?

# How many features (columns) are in the data?

# What, if any, ethical questions or questions of power do you need to consider when working with this data?

# What are possible limitations or problems with this data? (at least 200 words)

### Trends Over Time Chart

This chart shows black females incarcerated at disproportionately higher rates than the white female. Especially towards 2000 - 2010, there seems to be a peak in black female imprisonment and a large difference between white and black female imprisonment. Even within the female population, which is significantly less than the male population in prison, the disproportion between the black and white populations is significant. This is an important chart to include because it reveals the result of both historical and present-day racism. Furthermore, it visually shows the need to help eliminate the factors that contribute to this inequality. 
```{r chart_1, fig.width=12}
chart_1
```

### Variable Comparison Chart

In order to look more deeply into the disproportion between white and black females in prison, this chart visualizes the proportion in Texas, one of the highest total numbers of female imprisonment. From this chart, it is seen that there is a significant difference between the black and white female population in Texas prisons, with black females being the majority in most counties. 

```{r chart_2, fig.width=12}
chart_2
```

### Map

From this chart, it is clear that Texas has the largest number of black female prison populations, Florida and California coming in close seconds. The purpose of this chart is to see where the highest population of black female prisoners are. This can lead to the investigation of environmental factors that contribute to the higher rates. Furthermore, it reveals the question of whether the diversity of a state can lead to the increase or decrease of the black female prison population.

```{r prison_pop_map, fig.width=12}
prison_pop_map
```
