---
title: "A3: Incarceration"
author: Eun Be Cha
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(scales)
library(maps)

data_prison <- read.csv("https://github.com/melaniewalsh/Neat-Datasets/blob/main/us-prison-pop.csv?raw=true")

source('chart1.R')
source('chart2.R')
source('map.R')

```

### Introduction

I am choosing to evaluate the difference between the black female and white female prison population. Because America is the country with the greatest number of people incarcerated in the world and Black people are incarcerated at disproportionately high rates, I wanted to analyze this disproportion within the female population in prison. 

### Summary Information

The average male population in Arizona prison is 2161 people, compared to the female prison population of 208 people. The difference being 1953 people. However, most states seem to have a positive difference between the average white female and black female prison population, meaning there is a higher white female prison population. For instance, in Washington, 2015, the sum of white female prison population was 904 people, whereas, the black female prison population was 104. In 2015, the states with the most female black prison population were Texas, Florida, and California at 3498, 2091, and 1518 people respectively. Some states that have negative differences, meaning a higher black female prison population than white female population, include New York, New Jersey, Michigan, Georgia, and Maryland. 

### The Dataset

The data was released by the Vera Institute of Justice. They employed two Bureau of Justice Statistics (BJS) data collections: the Census of Jails (COJ) and the Annual Survey of Jails (ASJ). Furthermore, this project was funded by the Robert W. Wilson Charitable Trust. The data was collected to fill the gap of data needed to analyze the trends of causes, consequences, and inequalities. The data can be used to reveal trends such as environmental factors that increase or decreases a prison population group. Most of my analysis used the data from "Dataset about prison population per county/state." There are 153811 observations (rows) and 36 features (columns). Some features include the male and female population in jail and the population of different races. A possible limitation of the data set is that there are a lot of missing data (NA). There are also large gaps between the time taken between the data. This can make it harder to find accurate trends. Moreover, it is important to consider what else might be missing from the data such as data about the overall diversity and culture of the state or what the legal system is like for that specific state. A question of power that is worth considering when working with this data set is whether some of the NA in the data set is the result of hiding the statistics to conceal inequalities.

### Trends Over Time Chart

This chart shows black females incarcerated at disproportionately higher rates than the white female. Especially towards 2000 - 2010, there seems to be a peak in black female imprisonment and a large difference between white and black female imprisonment. Even within the female population, which is significantly less than the male population in prison, the disproportion between the black and white populations is significant. This is an important chart to include because it reveals the result of both historical and present-day racism. Furthermore, it visually shows the need to help eliminate the factors that contribute to this inequality. 
```{r chart_1, fig.width=12}
chart_1
```

### Variable Comparison Chart

In order to look more deeply into the disproportion between white and black females in prison, this chart visualizes the proportion in Texas, one of the highest total numbers of female imprisonment. From this chart, it is seen that there is a significant difference between the black and white female population in Texas prisons, with black females being the majority in most counties. 

```{r chart_2, fig.width=12}
chart_2
```

### Map

From this chart, it is clear that Texas has the largest number of black female prison populations, Florida and California coming in close seconds. The purpose of this chart is to see where the highest population of black female prisoners are. This can lead to the investigation of environmental factors that contribute to the higher rates. Furthermore, it reveals the question of whether the diversity of a state can lead to the increase or decrease of the black female prison population.

```{r prison_pop_map, fig.width=12}
prison_pop_map
```
